Supplementary Materials:\\France-wide 10\,m Deciduous–Evergreen Mapping: AlphaEarth Embeddings vs Harmonics

### 2025-10-16 article/supplementary_materials.tex
cool stuff

\subsection{Harmonic Feature Extraction}
\begin{enumerate}
    \item Sentinel-2 Level-2A time series extraction (2023, 10-day composites)
    \item Cloud masking using QA60 band and SCL classification
    \item Temporal interpolation and gap-filling
    \item Two-harmonic Fourier decomposition per vegetation index
    \item Feature aggregation at 10\,m pixel level
\end{enumerate}
Implementation: \texttt{src/features/extract\_harmonic\_features.py}

\subsection{Embedding Feature Extraction}
\begin{enumerate}
  \item Data source: Earth Engine “Satellite Embedding V1 (Annual)” (10\,m, 64D, 2017–2024) \citep{Google2025SatelliteEmbeddingV1}.
  \item Export year: 2023; exported at native 10\,m over forest AOIs; bands \texttt{A00..A63} (unit-normalized; public assets store int8 quantized values de-quantized on read).
  \item Processing: server-side clipping per tile; local mosaic to the training grid; no additional normalization.
  \item Note: some external work resamples embeddings to coarser grids (e.g., 500\,m) for continental-scale transfer; here we retain 10\,m for boundary fidelity \citep{Houriez2025AEFDataGen}.
\end{enumerate}
Implementation: \texttt{src/gee/} and \texttt{src/inference/} workflows.

@@

@@

- SatMAE pre-training for temporal and multi-spectral satellite imagery \citep{Cong2022}.
- Prithvi-EO multi-temporal geospatial foundation models and resources \citep{Szwarcman2024PrithviEO2}.
- Billion-scale remote sensing foundation model results across tasks \citep{Cha2023Billion}.
- Generative foundation modeling for satellite imagery \citep{Khanna2023DiffusionSat}.
- Discussion on when foundation models are effective for pixel-level classification \citep{Xie2024FoundationEffective}.

\section{S9. Product Comparisons (DLT, BD For\^et V2)}

We compare the 2023 harmonic-based phenology map against Copernicus DLT (broadleaved vs coniferous) and BD For\^et V2 (deciduous vs evergreen) using per-tile and per-eco-region summaries. Agreement is computed only where both products agree that a pixel is forest. National summary metrics are reported in Table~\ref{tab:product_comparison_national} of the main text.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/map_oa_2x2_DLT.png}
    \caption{Overall accuracy (2\,x\,2\,km tiles) between the harmonic phenology map and Copernicus DLT, evaluated where both maps indicate forest. Darker colors indicate higher agreement.}
    \label{fig:oa_map_dlt}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/map_oa_2x2_BDForet.png}
\caption{Overall accuracy (2\,x\,2\,km tiles) between the harmonic phenology map and BD For\^et V2, evaluated where both maps indicate forest.}
    \label{fig:oa_map_bdforet}
\end{figure}

@@

\begin{table}[H]
    \centering
\caption{Eco-region agreement: harmonic phenology vs DLT and BD For\^et V2. Values derived from results/comparison CSVs.}
    \begin{tabular}{lcccc}
        \hline
        Eco-region & OA vs DLT & Kappa vs DLT & OA vs BD For\^et & Kappa vs BD For\^et \\
        \hline
        Alps & 0.513 & 0.027 & 0.486 & -0.031 \\
        Central Massif & 0.588 & 0.092 & 0.569 & 0.077 \\
        Corsica & 0.288 & -0.008 & 0.462 & -0.011 \\
        Greater Crystalline and Oceanic West & 0.622 & 0.015 & 0.609 & 0.013 \\
        Greater Semi-Continental East & 0.787 & 0.074 & 0.772 & 0.084 \\
        Jura & 0.567 & -0.007 & 0.438 & -0.179 \\
        Mediterranean & 0.514 & 0.054 & 0.498 & -0.003 \\
        Oceanic Southwest & 0.689 & 0.368 & 0.698 & 0.347 \\
        Pyrenees & 0.584 & 0.060 & 0.617 & 0.145 \\
        Semi-Oceanic North Center & 0.774 & 0.044 & 0.771 & 0.024 \\
        Vosges & 0.515 & 0.029 & 0.367 & -0.327 \\
        \hline
    \end{tabular}
    \label{tab:product_comparison_ecoregion}
\end{table}

>>>>

\subsection{Embedding Feature Extraction}
\begin{enumerate}
    \item Data source: Google Earth Engine \texttt{Satellite Embedding V1 (Annual)} collection (10\,m, 64-D AlphaEarth vectors, 2017--2024) \citep{Google2025SatelliteEmbeddingV1}.
    \item For each study year (2018, 2020, 2022, 2023) we export the tiles at native 10\,m resolution, then clip/mosaic them to the 2.5\,km training grid.
    \item Bands \texttt{A00..A63} are stored as quantized int8 in Earth Engine; values are de-quantized to float32 on read before any feature selection.
    \item No additional normalization or smoothing is applied so that per-pixel embeddings remain comparable across years; downstream feature pruning (EMB-14) happens during cross-validation.
\end{enumerate}
Implementation: \texttt{src/gee/} and \texttt{src/inference/} workflows.

@@

\subsection{Calibration and Threshold Sensitivity}

We recomputed eco-region-balanced 5-fold cross-validation to extract out-of-fold probabilities for both Random Forest models. Key findings:

\begin{itemize}
    \item \textbf{Expected Calibration Error (ECE)}: Emb-Top14 = 0.033; Harm-Top14 = 0.059 (Maximum Calibration Error 0.114 vs 0.184).
    \item \textbf{Threshold sweep (0.45/0.50/0.55)}: Embedding macro-F1 spans 0.903–0.906; harmonic macro-F1 spans 0.870–0.879.
    \item \textbf{Fold variability}: Embedding macro-F1 standard deviation 0.003; harmonic 0.011.
\end{itemize}

Reliability histograms (10 bins) and per-threshold confusion counts are available in: \texttt{results/analysis\_calibration/embeddings/}\{\texttt{calibration\_metrics.json}, \texttt{reliability\_curve.csv}\} and \texttt{results/analysis\_calibration/harmonic/}\{\texttt{calibration\_metrics.json}, \texttt{reliability\_curve.csv}\}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.85\textwidth]{images/reliability_comparison.png}
    \caption{\textbf{Figure S3.} Reliability diagrams comparing Emb-Top14 (left) and Harm-Top14 (right) Random Forests. Bars show sample fraction per probability bin; circles plot mean confidence, squares plot empirical accuracy. The embedding model exhibits lower calibration error (ECE 0.033 vs 0.059) and smaller maximum deviation.}
    \label{fig:reliability}
\end{figure}

The per-eco-region accuracy/F1 deltas that back Table~\ref{tab:eco_region_performance} are stored in \texttt{results/analysis\_coherence/eco\_region\_delta\_metrics.csv} (generated via \texttt{src/analysis/compute\_eco\_region\_deltas.py}).

\subsection{Spatial Coherence Metrics}

To quantify the qualitative differences in map texture, we computed two coherence metrics per tile and aggregated nationally:
\begin{itemize}
    \item \textbf{Edge density} (m km$^{-2}$): total length of class boundaries divided by valid forest area;
    \item \textbf{Patch density} (per 100 km$^{2}$): number of 8-connected deciduous or evergreen patches scaled to 100 km$^{2}$.
\end{itemize}

Evaluated over 639 eco-balanced 2.5\,km tiles (sampling the national rasters):

\begin{table}[H]
    \centering
    \caption{\textbf{Table S3.} Spatial coherence (national means).}
    \begin{tabular}{lcc}
        \hline
        \textbf{Variant} & \textbf{Edge density (m km$^{-2}$)} & \textbf{Patch density (per 100 km$^{2}$)} \\ \hline
        Embedding (Top-14) & 3924 & 2925 \\
        Harmonic (Top-14) & 8789 & 8462 \\
        Harmonic + 3\,$\times$\,3 median & 4326 & 3150 \\ \hline
    \end{tabular}
\end{table}

Embeddings produce fewer boundaries and fewer isolated patches than harmonics, consistent with spatial context captured during pre-training. A simple 3\,$\times$\,3 median filter reduces harmonic speckle but embeddings remain ~9\% lower in edge density and ~7\% lower in patch density on average. Implementation and raw outputs are provided by the CLI \texttt{src/analysis/compute\_coherence\_metrics.py}, with summaries in \texttt{results/analysis\_coherence/coherence\_summary.csv}.

@@

\begin{itemize}
    \item SatMAE pre-training for temporal and multi-spectral satellite imagery \citep{Cong2022}.
    \item Prithvi-EO multi-temporal geospatial foundation models and resources \citep{Szwarcman2024PrithviEO2}.
    \item Billion-scale remote sensing foundation model results across tasks \citep{Cha2023Billion}.
    \item Generative foundation modeling for satellite imagery \citep{Khanna2023DiffusionSat}.
    \item Discussion on when foundation models are effective for pixel-level classification \citep{Xie2024FoundationEffective}.
\end{itemize}

\section{S9. Product Comparisons (DLT, BD For\^{e}t V2)}

We compare the 2023 harmonic-based phenology map against Copernicus DLT (broadleaved vs coniferous) and BD For\^{e}t V2 (deciduous vs evergreen) using per-tile and per-eco-region summaries. Agreement is computed only where both products agree that a pixel is forest. National summary metrics are reported in Table~\ref{tab:product_comparison_national} of the main text.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.49\textwidth]{images/oa2x2_embeddings_dlt.png}\hfill
    \includegraphics[width=0.49\textwidth]{images/oa2x2_embeddings_bdforet.png}\\[0.5em]
    \includegraphics[width=0.49\textwidth]{images/oa2x2_harmonic_dlt.png}\hfill
    \includegraphics[width=0.49\textwidth]{images/oa2x2_harmonic_bdforet.png}
    \caption{Tile-level agreement (OA, \%) computed where both maps indicate forest. Top: Embeddings vs DLT (left) and BD For\^{e}t (right). Bottom: Harmonics vs DLT (left) and BD For\^{e}t (right). Compact styling and percent colorbars highlight national patterns without clutter.}
    \label{fig:oa2x2_products}
\end{figure}

@@

The weighted eco-region summaries for both harmonic and embedding comparisons—versus DLT and BD For\^{e}t—are available in \texttt{results/comparison/embeddings\_vs\_dlt\_by\_region.csv}, \texttt{results/comparison/embeddings\_vs\_bdforet\_by\_region.csv}, and their harmonic counterparts. These tables include overall accuracy, kappa, and macro-F1 along with raw class-by-class confusion counts; analysts can slice them directly to build maps or statistics tailored to their regions of interest.

\section{S10. Multiclass Tree Taxa from Embeddings}

We trained Random Forests on the 64‑dimensional annual embeddings to classify tree taxa beyond deciduous–evergreen. At the genus level the model achieves overall accuracy (OA) of 0.828, weighted F1 of 0.829, and macro‑F1 of 0.387. Species‑level performance is lower but still competitive given the long‑tailed label distribution: OA 0.769, weighted F1 0.771, macro‑F1 0.289. Per‑eco‑region means reflect similar patterns (genus: OA 0.805, macro‑F1 0.381; species: OA 0.755, macro‑F1 0.278). These results indicate that embeddings capture taxonomic cues most strongly for dominant classes; tail performance is limited by class imbalance and label sparsity.

\subsection*{Eco‑region Performance}

\begin{table}[H]
    \centering
    \caption{\textbf{Table S4.} Genus classification by eco‑region.}
    \begin{tabular}{lcc}
        \hline
        Eco‑region & OA & Macro F1 \\
        \hline
        Semi-Oceanic North Center & 0.805 & 0.260 \\
        Unknown & 0.947 & 0.407 \\
        Mediterranean & 0.953 & 0.540 \\
        Corsica & 0.821 & 0.490 \\
        Alps & 0.863 & 0.478 \\
        Oceanic Southwest & 0.878 & 0.308 \\
        Greater Crystalline and Oceanic West & 0.697 & 0.317 \\
        Vosges & 0.616 & 0.366 \\
        Jura & 0.787 & 0.316 \\
        Pyrenees & 0.786 & 0.320 \\
        Greater Semi-Continental East & 0.745 & 0.380 \\
        Central Massif & 0.769 & 0.395 \\
        \hline
    \end{tabular}
    \label{tab:genus_ecoregion_embeddings}
\end{table}

\begin{table}[H]
    \centering
    \caption{\textbf{Table S5.} Species classification by eco‑region.}
    \begin{tabular}{lcc}
        \hline
        Eco‑region & OA & Macro F1 \\
        \hline
        Semi-Oceanic North Center & 0.763 & 0.232 \\
        Unknown & 0.897 & 0.521 \\
        Mediterranean & 0.834 & 0.275 \\
        Corsica & 0.746 & 0.297 \\
        Alps & 0.752 & 0.335 \\
        Oceanic Southwest & 0.822 & 0.200 \\
        Greater Crystalline and Oceanic West & 0.574 & 0.214 \\
        Vosges & 0.621 & 0.246 \\
        Jura & 0.783 & 0.237 \\
        Pyrenees & 0.771 & 0.189 \\
        Greater Semi-Continental East & 0.741 & 0.273 \\
        Central Massif & 0.758 & 0.319 \\
        \hline
    \end{tabular}
    \label{tab:species_ecoregion_embeddings}
\end{table}

\subsection*{Notes}
- Dominant taxa attain high F1 (often 0.64–0.90), while rare classes depress macro‑F1.
- Future improvements: class‑balanced sampling and losses, hierarchical genus→species modeling, eco‑region‑aware priors, probability calibration with top‑k reporting, and light spatial regularization.
